Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-25 robdml.tex
tex

%%%%%%%%% ABSTRACT
\begin{abstract}
	% background
	Owing to the security implications of adversarial vulnerability, the
	adversarial robustness of deep metric learning models have to be improved
	via defense.
	% insight
	In order to avoid model collapse due to excessively hard examples, the
	existing methods refrain from using the min-max adversarial training, but
	instead learns from a weak adversary inefficiently.
	% our starting point
	Conversely, we argue that a model can learn from a strong adversary without
	collapse, where the key is triplet hardness.
	% HM framework
	To this end, Hardness Manipulation is proposed to adversarially perturb the
	training triplet to increase its hardness into a specified level, according
	to another benign triplet or a pseudo-hardness function.
	% GA Adversary
	Based on this, Gradual Adversary is proposed as a family of pseudo-hardness
	functions to overcome problems of referencing other benign triplets, and
	achieve a better balance in the learning objective.
	% Intra-Class Structure
	Besides, a penalty on Intra-Class Structure among benign and adversarial
	examples can further improve model robustness, which is neglected by
	previous methods.
	% Experiment and Conclusion
	Comprehensive experimental results on commonly used deep metric
	learning datasets suggest that the proposed method, although simple in
	form, overwhelmingly outperforms the state-of-the-art defense methods in
	terms of training time cost, robustness, as well as recall performance on
	benign examples.
%
\end{abstract}

@@

Existing defense methods for DML~\cite{advrank,robrank} are adversarial
training-based, inspired by Madry's \emph{min-max} adversarial
training~\cite{madry} because it is one of the most effective methods for deep
neural network classifiers.
%
Specifically, Madry's method involves a inner problem to \emph{maximize} the
loss by perturbing the inputs into adversarial examples, and an outer problem
to \emph{minimize} the loss by updating the model parameters.
%
However, in order to avoid model collapse, the existing DML defenses refrain
from directly adopting such paradigm, but instead use some indirect and
relatively weak objectives to merely increase the loss value to a certain
level.
%
As a result, the model is not learning from a strong adversary, and suffers
from low efficiency in gaining robustness within limited time. 
%
Furthermore, adversarial training already suffers from significant training
time cost (compared to the normal training process).
%
Hence, time efficiency in gaining adversarial robustness
is an inevitable major issue.

@@

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified \emph{destination}
hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
\emph{destination} hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified \emph{destination} hardness level.
%
But, how hard should such ``\emph{destination} hardness'' be?

@@

Recall that the model is already prone to collapse with excessively hard benign
triplets, let alone adversarial examples.
%
Thus, intuitively, the \emph{destination} hardness can be that of another
benign triplet which is harder than the given triplet (\eg, a
Semihard~\cite{facenet} triplet).
%
However, in the late phase of training, the expectation of the loss, as well as
the corresponding \emph{hardness} of a given triplet will be small, which means
the \emph{destination} hardness will be close to that of the sample triplet for
training.
%
This leads to weak attacks, and inefficient learning from a weak adversary.
%
Besides, we speculate a strong adversary in the early phase of DML training may
hinder the model from learning good embeddings, and hence influence the
eventual performance on benign examples.
%
Based on these, a \emph{destination} hardness calculated from benign examples
may not be the best choice.
%
A better \emph{destination} hardness should be able to balance the training
objectives in the early and late phases of training.

@@

To this end, we propose Graduate Adversary, a family of pseudo-hardness
functions which can be used as the \emph{destination} hardness in HM.
%
Any function that leads to a relatively weak adversary in the early training
phase, and a relatively strong adversary in the late training phase belongs to
this family.
%
As an example, we design a ``Simple Graduate Adversary'' (SGA) pseudo-hardness
function as the linearly scaled negative triplet margin, incorporating a strong
assumption that the \emph{destination} hardness should remain Semihard.
%
Empirical observation will show the advantage of such pseudo-hardness
functions.

@@

\textbf{Adversarial Defense.}
%
To battle against the attacks, some early defenses incur gradient masking
effect, but it gives a false sense of security~\cite{obfuscated}.
%
Defensive distillation~\cite{distill2} can be compromised~\cite{cw}.
%
Ensemble of weak defenses is not robust~\cite{ensembleweak}.
%
Other forms of defenses are also proposed, such as input
preprocessing~\cite{deflecting}, or a randomization inside the
network~\cite{self-ensemble}.
%
But various defense methods are still susceptible to adaptive
attack~\cite{adaptive}.
%
Of all defenses, adversarial training remains to be one of the most
effective method~\cite{bilateral,advtrain-triplet,benchmarking}, and it
consistently show promising empirical robustness,
although suffering from significant training time cost and a lower performance
on benign examples.

@@

The model performance on benign (\ie, unperturbed) examples is measured in
Recall@1 (R@1), Recall@2 (R@2), mAP and NMI
following~\cite{revisiting,robrank}.
%
The adversarial robustness of a model is measured in Empirical Robustness Score
(ERS)~\cite{robrank}, a normalized score (the higher the better) from
an ensemble of existing attacks against DML or their simplified white-box
alternatives.
%
Since adversarial training is not gradient masking~\cite{obfuscated}, the
performance of white-box attack can be regarded as the upper bound of the
black-box attacks, and thus a model that is empirically robust to the ensemble
of white-box attacks is expected to be robust in general.

\oo{Concretely, the ensemble of attack for computing ERS includes.}
%
Further details can be found in the supplementary.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
	% background
	Owing to the security implications of adversarial vulnerability, the
	adversarial robustness of deep metric learning models has to be improved
	via defense.
	% insight
	In order to avoid model collapse due to excessively hard examples, the
	existing defenses refrain from using the min-max adversarial training, but
	instead learns from a weak adversary inefficiently.
	% our starting point
	Conversely, we argue that a model can learn from a strong adversary without
	collapse, where the key is triplet hardness.
	% HM framework
	To this end, Hardness Manipulation is proposed to adversarially perturb the
	training triplet to increase its hardness into a specified level, according
	to another benign triplet or a pseudo-hardness function.
	% GA Adversary
	Based on this, Gradual Adversary is proposed as a family of pseudo-hardness
	functions to overcome problems of referencing other benign triplets, and
	achieve a better balance in the learning objective.
	% Intra-Class Structure
	Besides, a penalty on Intra-Class Structure among benign and adversarial
	examples can further improve model robustness, which is neglected by
	previous methods.
	% Experiment and Conclusion
	Comprehensive experimental results on commonly used deep metric
	learning datasets suggest that the proposed method, although simple in
	its form, overwhelmingly outperforms the state-of-the-art defense methods in
	terms of training time cost, robustness, as well as recall performance on
	benign examples.
%
\end{abstract}

@@

Existing defense methods for DML~\cite{advrank,robrank} are adversarial
training-based, inspired by Madry's \emph{min-max} adversarial
training~\cite{madry} because it is one of the most effective methods for deep
neural network classifiers.
%
Specifically, Madry's method involves a inner problem to \emph{maximize} the
loss by perturbing the inputs into adversarial examples, and an outer problem
to \emph{minimize} the loss by updating the model parameters.
%
However, in order to avoid model collapse, the existing DML defenses refrain
from directly adopting such paradigm, but instead use some indirect and
relatively weak objectives to increase the loss value to a certain level.
%
As a result, the model is not learning from a strong adversary, and suffers
from low efficiency in gaining robustness within limited time. 
%
Furthermore, adversarial training already suffers from significant training
cost (compared to the normal training process).
%
Hence, efficiency in gaining adversarial robustness is an inevitable major
issue for DML defense methods.

@@

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified \emph{destination}
hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
\emph{destination} hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified \emph{destination} hardness level.

@@

Then, how hard should such ``\emph{destination} hardness'' be?
%
Recall that the model is already prone to collapse with excessively hard benign
triplets, let alone adversarial examples.
%
Thus, intuitively, the \emph{destination} hardness can be that of another
benign triplet which is harder than the given triplet (\eg, a
Semihard~\cite{facenet} triplet).
%
However, in the late phase of training, the
expectation of \emph{destination} hardness of another benign triplet
will be small, and close to the \emph{hardness} expectation of the given triplet.
%
This leads to weak attacks, and inefficient learning from a weak adversary.
%
Besides, we speculate a strong adversary in the early phase of DML training may
hinder the model from learning good embeddings, and hence influence the
eventual performance on benign examples.
%
Based on these, a \emph{destination} hardness calculated from benign examples
may not be the best choice.
%
A better \emph{destination} hardness should be able to balance the training
objectives in the early and late phases of training.

@@

To this end, we propose Graduate Adversary, a family of pseudo-hardness
functions which can be used as the \emph{destination} hardness in HM.
%
Any function that leads to a relatively weak adversary in the early training
phase, and a relatively strong adversary in the late training phase belongs to
this family.
%
As an example, we design a ``Simple Graduate Adversary'' (SGA) pseudo-hardness
function as the linearly scaled negative triplet margin, incorporating a strong
assumption that the \emph{destination} hardness should remain Semihard.
%
%Empirical observation will show the advantage of such pseudo-hardness
%functions.

% Intra-Class Structure

Besides, a sample triplet will be augmented into a sextuplet (benign and
adversarial examples) during adversarial training, where a \emph{intra-class}
structure can be enforced.
%
Such structure has been neglected by existing methods, but some existing
attacks aims to change the sample rankings in the same class~\cite{advrank}.
%
Hence, we propose a simple \emph{intra-class} structure penalty term for
adversarial training, which is expected to improve robustness.

@@

\textbf{Adversarial Defense.}
%
To battle against the attacks, some early defenses incur gradient masking
effect, but it gives a false sense of security~\cite{obfuscated}.
%
Defensive distillation~\cite{distill2} can be compromised~\cite{cw}.
%
Ensemble of weak defenses is not robust~\cite{ensembleweak}.
%
Other forms of defenses are also proposed, such as input
preprocessing~\cite{deflecting}, or a randomization inside the
network~\cite{self-ensemble}.
%
But various defense methods are still susceptible to adaptive
attack~\cite{adaptive}.
%
Of all defenses, adversarial training remains to be one of the most
effective method~\cite{bilateral,advtrain-triplet,benchmarking}, and it
consistently show promising empirical robustness,
although suffering from significant training cost and a lower performance
on benign examples.

@@

The model performance on benign (\ie, unperturbed) examples is measured in
Recall@1 (R@1), Recall@2 (R@2), mAP and NMI
following~\cite{revisiting,robrank}.
%
The adversarial robustness of a model is measured in Empirical Robustness Score
(ERS)~\cite{robrank}, a normalized score (the higher the better) from
an collection of existing attacks against DML or their simplified white-box
alternatives.
%
Since adversarial training is not gradient masking~\cite{obfuscated}, the
performance of white-box attack can be regarded as the upper bound of the
black-box attacks, and thus a model that is empirically robust to the collection
of white-box attacks is expected to be robust in general.

\oo{Concretely, the collection of attack for computing ERS includes.}
%
Further details can be found in the supplementary.
